---
title: "RC1_Linear_Regression"
author: "Haohong Shang"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This slide is created by UM's GSI.

## Import dataset

We'll work with the Auto dataset from the Lab 1 exercise

```{r}
library(ISLR)

data(Auto)
# what are the variables?
names(Auto)

```

The ISLR package provides a more detailed description if you type ?Auto in the console.

We'll begin by fitting a simple linear regression to predict mpg given horsepower using the lm function.

## Fitting simple linear regression with lm

```{r}
attach(Auto)

# basic syntax is lm(y~x), which produces a linear model object
lm.fit <- lm(mpg~horsepower)

```

## Inspect fitted model

We can see basic output by printing lm.fit. More detailed output is given by calling summary(lm.fit), incluing R-squared, and standard errors of coefficients.

You should make yourself familiar with all the information in the summary output.

```{r}
# basic lm output
lm.fit

# detailed lm output
summary(lm.fit)

```

lm automatically produces some other objects we may be interested in (e.g. vector of residuals, vector of coefficients).

```{r}

# view the available lm objects
names(lm.fit)

# there are two ways to access the coefficient vector, for instance
# (1)
lm.fit$coefficients
# (2)
coef(lm.fit)
```

R will automatically provide confidence intervals for the parameters of the fitted model, and predictions for new data

```{r}

# confidence intervals for coefficient vector
confint(lm.fit)

# predictions for new data (with error bounds)
predict(lm.fit,data.frame(horsepower=c(100,120,200)),interval="prediction")

```

## Plots and visual diagnostics

We can plot the fitted regression line with the abline function

```{r}
# scatter plot of horsepower and mpg
plot(horsepower,mpg,main="Simple regression line")
# add regression line
abline(lm.fit,col="red")

```

R also produces four basic diagnostic plots with plot.lm:

* Residuals vs fitted values
* Residual QQ-plot
* Standardized residuals vs fitted values
* Standardized resiudals vs leverage

```{r}
# set plot options so all four plots appear in the same pane
par(mfrow=c(2,2))

# diagnostic plots
plot(lm.fit)

# reset plot options
par(mfrow=c(1,1))

```

We can produce these plots ourselves by using the objects produced by lm, for instance suppose we want to see the residuals vs fitted values plot.

```{r}
# residuals v fitted values plot
y.hat <- predict(lm.fit)
e.hat <- residuals(lm.fit)
plot(y.hat,e.hat,
     xlab="Fitted values",ylab="Residuals",
     main="Residuals v Fitted values")
abline(h=0,lty=3)
```

## Fitting multiple regression with lm 

Suppose we also think vehicle weight will influence fuel efficiency. To fit a multiple linear regression, simply add more variables to the formula in lm.

```{r}
# fit a multiple linear regression
mlm.fit <- lm(mpg~horsepower+weight)

# view summary output
summary(mlm.fit)

```

A few things to note:

* Both predictors appear to have an influence on fuel efficiency
* The R-squared value has increased after adding another predictor
* The estimate of the coefficient for horsepower has changed

Suppose we want to add even more parameters to the model. The formula shortcut to add all variables is '.' (**Warning**: the '.' command is not compatible with 'attach', you'll have to specify the dataset in the lm call)

However in this case, the data contains non-numeric variables 'origin' and 'name' which we we can remove from the model by subtracting them in the formula.

```{r}
# fit a multiple linear regression
mlm.fit.2 <- lm(mpg~.-origin-name,data=Auto)

# view summary output
summary(mlm.fit.2)

```

## Fitting regression with transformed predictors with lm

Recall the plot we produced of residuals vs fitted values in our simple linear regression (y=mpg, x=horsepower). We can see a clear quadratic pattern in this plot, implying we could improve the fit by adding a quadratic term to our model.

There are three ways to do this:

(1) Create a new column in the data frame
```{r}
# create a column containing the squared horsepower
Auto$horsepower2 <- horsepower^2

# fit a multiple linear regression on horsepower and horsepower^2
qm.fit <- lm(mpg~horsepower + horsepower2,data=Auto)

# view basic output
qm.fit

```

(2) Do the transformation directly in lm
```{r}
# fit a multiple linear regression
qm.fit2 <- lm(mpg~horsepower+I(horsepower^2))

# view basic output
qm.fit2

```

(3) Use the poly command to fit a higher order polynomial model
```{r}
# fit a multiple linear regression
qm.fit3 <- lm(mpg~poly(horsepower,2))

# view basic output
qm.fit3

# the poly command uses orthonormal polynomials, so it produces 
# different coefficients but the same fitted values
plot(fitted.values(qm.fit),fitted.values(qm.fit3),
     main="Comparison of fitted values")
abline(a=0,b=1,lty=2)

```

The poly command can be used to fit even higher order polynomial models.

After adding a quadratic term, we have removed the pattern in the residual vs fitted value plot. 

```{r}
# quadratic model residuals v fitted values plot
y.hat <- predict(qm.fit)
e.hat <- residuals(qm.fit)
plot(y.hat,e.hat,
     xlab="Fitted values",ylab="Residuals",
     main="Residuals v Fitted values (quadratic model)")
abline(h=0,lty=3)
```


